Demo arXiv template

### 2022-10-04 paper/paper.qmd
add more content

---
title: "Demo arXiv template"
format:
  arxiv-pdf:
    keep-tex: true  
  arxiv-html: default
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
    url: https://huizezhangsh.netlify.app/
  - name: Collaborators
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
abstract: |
  - The paper describes a framework for constructing indices 
  - use drought indices as examples, but appliable in general to environmental indices constructed from multivariate spatio-temporal data
keywords: 
  - spatio-temporal data
  - indices
  - data pipeline
bibliography: bibliography.bib  
---

@@

Various methods are proposed to extract multivariate information on the spatial extent, across time.

@@

@@

@@

@@

only for distribution fit, convert from an fitted distribution to normal

@@

>>>>

---
title: "Demo arXiv template"
format:
  arxiv-pdf:
    keep-tex: true  
  arxiv-html: default
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@monash.edu
    url: https://huizezhangsh.netlify.app/
  - name: Collaborators
    affiliations:
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, VIC
        country: Australia
abstract: |
  - Indices, useful, quantify severity, early monitoring, 
  - A huge number of indices have been proposed by domain experts, however, a large majority of them are not being adopted and reused in research or in practice. 
  - One of the most commonly used drought index, SPI and SPEI, are available through the `SPEI` package and its readily avaiability to be applied on user's data 
  - However, even if most of the indices have available existing implementation available to users, there would be hundreds of them and it will be burdensome to remember
  - The paper describes a general pipeline framework to construct indices from spatio-temporal data, so that indices can be constructed through [a set of general steps].
  - The benefit of such a framework is that ...
  - The framework will be demonstrated with drought indices as examples, but appliable in general to environmental indices constructed from multivariate spatio-temporal data
keywords: 
  - spatio-temporal data
  - indices
  - data pipeline
bibliography: bibliography.bib  
---

@@

Various methods are proposed to extract multivariate information on the spatial coverage, across time.

Indices with readily available code to implement have more potential to be computed by researchers, i.e. the package `SPEI` [@SPEI] has implemented the SPI and SPEI index for calculating drought indices. [Another package in a different domain]

@@

## Economics/ social indices

@@

## Tidy data

Before the concept of tidy data [@wickham_tidy_2014], tabular data arrive at data analysts in all different ways. 
Different analysts would write customised scripts for analysing the specific data. 
These scripts can be extended to other data analysed by the same people or group but this is not generalizable directly to another dataset. 
When the tidy data concept comes, variables and values are arranged so that 1) Each variable forms a column, 2) Each observation forms a row, and; 3) Each type of observational unit forms a table.
With this specific layout, wrangling on tabular data can be standardised into a grammar of data manipulation in `dplyr` [@dplyr].

A similar issue happens with index construction where researchers construct their own indices in their own ways and there has not yet been a tidy principle on index construction.
Also, this tidy principle on index construction is more complex than those in tidy data and the `dplyr` package.
It has to encompass the workflow of transformation from the raw data towards the final index series.

## Data pipeline

@@

The most popular indices (i.e. SPI, SPEI, etc) have existing software implementation (`SPEI`) to be applied to a different set of data.

constructing time series index should also be encapsulated in my framework

reusable

@@

specifically for converting from a fitted distribution to normal score via reverse CDF function, non-parametric formula, or empirical approximation, a common step in many index: SPI, SSI, Z score. The purpose of normalising is to convert the index into a standardised series after all the steps for the ease of comparison.

@@

Discretise the continuous index into a few labelled categories. For communicating the severity of natural hazard to general public.

### 2022-10-04 paper/paper.tex
add more content

\title{Demo arXiv template}
\author{
\textbf{H. Sherry Zhang}~\orcidlink{0000-0002-7122-1463}\\Department of
Econometrics and Business Statistics\\Monash University\\Melbourne,
VIC\\\href{mailto:huize.zhang@monash.edu}{huize.zhang@monash.edu}\\\\\\
\textbf{Collaborators}\\Department of Econometrics and Business
Statistics\\Monash University\\Melbourne, VIC\\}
\date{}
\begin{document}
\maketitle
\begin{abstract}
\begin{itemize}
\tightlist
\item
  The paper describes a framework for constructing indices
\item
  use drought indices as examples, but appliable in general to
  environmental indices constructed from multivariate spatio-temporal
  data
\end{itemize}
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
spatio-temporal data \sep indices \sep 
data pipeline

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, breakable, sharp corners, frame hidden]}{\end{tcolorbox}}\fi

@@

Numerous indices have been proposed by researchers and practitioners to
monitor natural hazard, for example, Alahacoon and Edirisinghe (2022)
reviews 111 drought indices derived from traditional and remote sensing
data; {[}review of index construction in other area, climate indices,
many other reviews in drought{]}.

Various methods are proposed to extract multivariate information on the
spatial extent, across time.

@@

@@

@@

The data pipeline in interactive graphics is a set of steps that
transform the raw data to the plots displayed on the screen. The initial
pipeline proposed by Buja et al. (1988) involves the following steps:
non-linear transformation, variables standardization, randomization,
projection engine, and viewporting. Another example in the early work of
pipeline by Sutherland et al. (2000) describes a three-step pipeline:
variable standardization, dimension reduction, and scaling data into the
viewing window. This pipeline also includes the transformation on
spatial and temporal variables, i.e.~computing time lag on temporal
variables. Wickham et al. (2009) argues that whether made explicit or
not, pipeline has to be presented in every graphics program and breaking
down graphic rendering into steps is also beneficial for understand the
implementation and compare between different graphic systems.

@@

@@

@@

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-alahacoon_comprehensive_2022}{}}%
Alahacoon, Niranga, and Mahesh Edirisinghe. 2022. {``A Comprehensive
Assessment of Remote Sensing and Traditional Based Drought Monitoring
Indices at Global and Regional Scale.''} \emph{Geomatics, Natural
Hazards and Risk} 13 (December): 762--99.
\url{https://doi.org/10.1080/19475705.2022.2044394}.

@@

@@

>>>>

\title{Demo arXiv template}
\author{
\textbf{H. Sherry Zhang}~\orcidlink{0000-0002-7122-1463}\\Department of
Econometrics and Business Statistics\\Monash University\\Melbourne,
VIC\\\href{mailto:huize.zhang@monash.edu}{huize.zhang@monash.edu}\\\\\\
\textbf{Collaborators}\\Department of Econometrics and Business
Statistics\\Monash University\\Melbourne, VIC\\}
\date{}
\begin{document}
\maketitle
\begin{abstract}
\begin{itemize}
\tightlist
\item
  Indices, useful, quantify severity, early monitoring,
\item
  A huge number of indices have been proposed by domain experts,
  however, a large majority of them are not being adopted and reused in
  research or in practice.
\item
  One of the most commonly used drought index, SPI and SPEI, are
  available through the \texttt{SPEI} package and its readily
  avaiability to be applied on user's data
\item
  However, even if most of the indices have available existing
  implementation available to users, there would be hundreds of them and
  it will be burdensome to remember
\item
  The paper describes a general pipeline framework to construct indices
  from spatio-temporal data, so that indices can be constructed through
  {[}a set of general steps{]}.
\item
  The benefit of such a framework is that \ldots{}
\item
  The framework will be demonstrated with drought indices as examples,
  but appliable in general to environmental indices constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
{\bfseries \emph Keywords}
\def\sep{\textbullet\ }
spatio-temporal data \sep indices \sep 
data pipeline

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, sharp corners, frame hidden, interior hidden]}{\end{tcolorbox}}\fi

@@

Various methods are proposed to extract multivariate information on the
spatial coverage, across time.

Indices with readily available code to implement have more potential to
be computed by researchers, i.e.~the package \texttt{SPEI} (Beguería and
Vicente-Serrano 2017) has implemented the SPI and SPEI index for
calculating drought indices. {[}Another package in a different domain{]}

@@

\hypertarget{economics-social-indices}{%
\subsection{Economics/ social indices}\label{economics-social-indices}}

@@

\hypertarget{tidy-data}{%
\subsection{Tidy data}\label{tidy-data}}

Before the concept of tidy data (Wickham 2014), tabular data arrive at
data analysts in all different ways. Different analysts would write
customised scripts for analysing the specific data. These scripts can be
extended to other data analysed by the same people or group but this is
not generalizable directly to another dataset. When the tidy data
concept comes, variables and values are arranged so that 1) Each
variable forms a column, 2) Each observation forms a row, and; 3) Each
type of observational unit forms a table. With this specific layout,
wrangling on tabular data can be standardised into a grammar of data
manipulation in \texttt{dplyr} (Wickham et al. 2022).

A similar issue happens with index construction where researchers
construct their own indices in their own ways and there has not yet been
a tidy principle on index construction. Also, this tidy principle on
index construction is more complex than those in tidy data and the
\texttt{dplyr} package. It has to encompass the workflow of
transformation from the raw data towards the final index series.

\hypertarget{data-pipeline}{%
\subsection{Data pipeline}\label{data-pipeline}}

@@

The data pipeline in interactive graphics is a set of steps that
transform the raw data to the plots displayed on the screen. The initial
pipeline proposed by Buja et al. (1988) involves the following steps:
non-linear transformation, variables standardization, randomization,
projection engine, and viewporting. The initial pipeline proposed by
Buja et al. (1988) involves the following steps: non-linear
transformation, variables standardization, randomization, projection
engine, and viewporting. Another example in the early work of pipeline
by Sutherland et al. (2000) describes a three-step pipeline: variable
standardization, dimension reduction, and scaling data into the viewing
window. This pipeline also includes the transformation on spatial and
temporal variables, i.e.~computing time lag on temporal variables. This
pipeline also includes the transformation on spatial and temporal
variables, i.e.~computing time lag on temporal variables. Wickham et al.
(2009) argues that whether made explicit or not, pipeline has to be
presented in every graphics program and breaking down graphic rendering
into steps is also beneficial for understand the implementation and
compare between different graphic systems.

@@

The most popular indices (i.e.~SPI, SPEI, etc) have existing software
implementation (\texttt{SPEI}) to be applied to a different set of data.

constructing time series index should also be encapsulated in my
framework

reusable

@@

Take one or multiple variables to create a new variable, linearly?
sometimes called feature extraction in the machine learning community
With drought indices, the extraction of meaningful variables from the
original data is usually supported by the water balance model, for
example, in SPEI, the step that create \(d\) out of precipitation and
potential evapotranspiration (PET) has theoretical backup from {[}see
paper.{]}

@@

specifically for converting from a fitted distribution to normal score
via reverse CDF function, non-parametric formula, or empirical
approximation, a common step in many index: SPI, SSI, Z score. The
purpose of normalising is to convert the index into a standardised
series after all the steps for the ease of comparison.

@@

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-SPEI}{}}%
Beguería, Santiago, and Sergio M. Vicente-Serrano. 2017. \emph{SPEI:
Calculation of the Standardised Precipitation-Evapotranspiration Index}.
\url{https://CRAN.R-project.org/package=SPEI}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-wickham_tidy_2014}{}}%
Wickham, Hadley. 2014. {``Tidy {Data}.''} \emph{Journal of Statistical
Software} 59 (September): 1--23.
\url{https://doi.org/10.18637/jss.v059.i10}.

@@

\leavevmode\vadjust pre{\hypertarget{ref-dplyr}{}}%
Wickham, Hadley, Romain François, Lionel Henry, and Kirill Müller. 2022.
\emph{Dplyr: A Grammar of Data Manipulation}.
